Effects

### 2020-01-07 doc/Effects.Rmd
close to wrapping up Effect summaries

Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.

```{r female.age, eval=FALSE}
#compile this dataset
```

@@

Check if there is a maternal age effect for females and males.

```{r mat.age.effect, echo=FALSE, eval=FALSE}
#pause evaluation -- needs debugging
meta.data.file = read.csv("~./MLH1repo/data/DandM_8.30.19.csv", header = TRUE)

meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
#445? -- probably extras at bottom not good

#how many mice have maternal DOB
meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
#151

@@

#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")

#str(meta.data.file.w.mat.DOB)

@@

#merge meta data with MLH1 -- rename this
see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
#remove anything without mean_CO
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
#what did I do with this DF...?

@@

#hist of data
mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
mat.age.hist.female

>>>>

Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.
Note all strains have both embryo and neonate samples. Subset for just, WSB, G, LEW, and PWD.

```{r female.age.setup, echo=FALSE}
#compile this dataset
#all female -- meta.dissections has emb or neonate?
#which sheet has this informaiton?

#clean.Meta.Data.txt
clean.meta.data <- read.csv("~./MLH1repo/data/clean.Meta.Data.txt", header = TRUE, sep='\t')
female.age.DF <- clean.meta.data[clean.meta.data$sex == "female",]#166

mergeFULL_F.mlh1 <- merge(female.age.DF, MLH1_data, by = 'mouse')

merge_F.mlh1_cell <- mergeFULL_F.mlh1[ ( (mergeFULL_F.mlh1$strain.x == 'WSB') |
                                       (mergeFULL_F.mlh1$strain.x == 'G') |
                                       (mergeFULL_F.mlh1$strain.x == 'LEW') |
                                  (mergeFULL_F.mlh1$strain.x == 'PWD') ), ]#901

#mouse level
mergeFULL_F.mlh1_mouse <- merge(female.age.DF, AP_mouse_table.HQ, by = 'mouse')#69 mice

merge_F.mlh1_mouse <- mergeFULL_F.mlh1_mouse[ ( (mergeFULL_F.mlh1_mouse$strain.x == 'WSB') |
                                       (mergeFULL_F.mlh1_mouse$strain.x == 'G') |
                                       (mergeFULL_F.mlh1_mouse$strain.x == 'LEW') |
                                  (mergeFULL_F.mlh1_mouse$strain.x == 'PWD') ), ]#42

```{r female.age.test_cells, echo=FALSE}
#logistic regression for MLH1 ~ emb | neonate

#code emb and neonate to 0,1
merge_F.mlh1_cell$age.stage <- ifelse(grepl("emb", merge_F.mlh1_cell$female.stage), 0, 
                       ifelse(grepl("neonate",merge_F.mlh1_cell$female.stage), 1, "other"))
merge_F.mlh1_cell$age.stage <- as.numeric(merge_F.mlh1_cell$age.stage)

logr_age.stage.CELLS <- glm(age.stage ~ nMLH1.foci, data=merge_F.mlh1_cell[merge_F.mlh1_cell$strain.x == "WSB",], family=binomial(link="logit"))
summary(logr_age.stage.CELLS)$coefficients

#
plot.blah.Cells <- ggplot(merge_F.mlh1_cell, aes(x=nMLH1.foci, y=age.stage)) + geom_point() +   stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Cell counts Female age logistic regression \nemb=0, neonate =1")+facet_wrap(~strain.x)
plot.blah.Cells

For the cell level curves  (all strains pooled) the p value is not significant. When I isolate the WSB observations, the p value is 0.0778.

```{r female.age.test_mouse, echo=FALSE}

merge_F.mlh1_mouse$age.stage <- ifelse(grepl("emb", merge_F.mlh1_mouse$female.stage), 0, 
                       ifelse(grepl("neonate",merge_F.mlh1_mouse$female.stage), 1, "other"))
merge_F.mlh1_mouse$age.stage <- as.numeric(merge_F.mlh1_mouse$age.stage)

#model fomula
logr_age.stage.MOUSE <- glm(age.stage ~ mean_co, data=merge_F.mlh1_mouse, family=binomial(link="logit"))
#logr_age.stage.MOUSE <- glm(age.stage ~ mean_co, data=merge_F.mlh1_mouse[merge_F.mlh1_mouse$strain.x == "WSB",], family=binomial(link="logit"))
summary(logr_age.stage.MOUSE)$coefficients

plot.blah.MOUSE <- ggplot(merge_F.mlh1_mouse, aes(x=mean_co, y=age.stage)) + geom_point() +   stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse Female age logistic regression \nemb=0, neonate = 1")+facet_wrap(~strain.x)
plot.blah.MOUSE
```

The mouse curves are more slanted, but the logistic regressions are not significant. The lg for WSB mouse averages isn't significant.

@@

Check if there is a maternal age effect for females and males.
clea.Meta.Data doesn't have maternal age data.

Collect data, get a decent set / large set of mice with maternal ages (mostly want to look at females).

```{r mat.age.effect, echo=FALSE}
#pause evaluation -- needs debugging

#this would be the best one so far
meta.data.file_FULL = read.csv("C:/Users/alpeterson7/Documents/MLH1repo/data/DandM_8.30.19.csv", header = TRUE)#updated with a couple more

#are there duplicate mice names?
mat.age_dupies <- meta.data.file_FULL[duplicated(meta.data.file_FULL$mouse),]
#these are blank mouse things

#try updating the D and M document, #clean up the data
#only those with mat age
meta.data.file_mat.age <- meta.data.file_FULL[!(is.na(meta.data.file_FULL$maternal_DOB)|meta.data.file_FULL$maternal_DOB==""),]
#meta.data.file <- droplevels(meta.data.file)

#needs mouse DOB
meta.data.file_mat.age <- meta.data.file_mat.age[!(is.na(meta.data.file_mat.age$DOB)|meta.data.file_mat.age$DOB==""),]#140

#check the date formats 
meta.data.file_mat.age$mat_DOB.for <- as.Date(meta.data.file_mat.age$maternal_DOB, format= "%m/%d/%Y")
#length(meta.data.file_mat.age$mouse)#152

meta.data.file_mat.age$maternal_DOB <- as.Date(meta.data.file_mat.age$maternal_DOB, format= "%m/%d/%Y")
meta.data.file_mat.age$DOB <- as.Date(meta.data.file_mat.age$DOB, format= "%m/%d/%Y")

#mat.age = euth date - mat.DOB
meta.data.file_mat.age <- add_mat_age(meta.data.file_mat.age)
#DOB, maternal_DOB .. got the

#merge this mat age file with mlh1 and mouse means!!
mergeFULL_mat.age_cells <- merge(meta.data.file_mat.age, MLH1_data, by = 'mouse')#952 (fewer than last time)

mergeFULL_mat.age_mouse <- merge(meta.data.file_mat.age, AP_mouse_table.HQ, by = 'mouse')#41

#now look at signal / pattern / #hist of data
mat.age.hist.plot <- ggplot(mergeFULL_mat.age_mouse, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")

mat.age.hist.plot_cells <- ggplot(mergeFULL_mat.age_cells, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females cells")

#plot with MLH1 data
mat.age_MLH1_cells <- ggplot(mergeFULL_mat.age_cells, aes(x = mat.age.days, y=nMLH1.foci, color = mouse))+geom_jitter(alpha=.3, width = 2)+facet_wrap(sex~subsp)+ggtitle("Maternal ages and MLH1")+theme(legend.position = 'none')

mat.age_MLH1_cells

mat.age_MLH1_mouse <- ggplot(mergeFULL_mat.age_mouse, aes(x = mat.age.days, y=mean_co, color = mouse))+geom_jitter(alpha=.3, width = 2)+facet_wrap(sex~subsp)+ggtitle("Maternal ages and MLH1")+theme(legend.position = 'none')

mat.age_MLH1_mouse

#mat.age_MLH1_cells.KAZ <- ggplot(mergeFULL_mat.age_cells[mergeFULL_mat.age_cells$category == "KAZ male",], aes(x = mat.age.days, y=nMLH1.foci, color = mouse))+geom_jitter()+facet_wrap(~category)+ggtitle("subset mice for maternal ages")+theme(legend.position = 'none')
#mat.age_MLH1_cells.KAZ
```

there should be plots above this
Based on the histograms of ages and the range across categories isn't very wide -- so not sure what the best test is.

```{r mat.format.rand, echo=FALSE, eval=FALSE}

@@

#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4

@@

@@
